HHHOn the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-04-05 TTau_Halpha/revision01/response_to_ref.txt

Dear Referee,

Thank you very much for reviewing our paper, and for doing so in
timely mannar.  We have reviewed our paper by considering your
comments and suggestions very carefully.  
The answers to your questions follows "*" marks before 
the indented texts below.

Kind regards,

Ryuichi Kurosawa, Tim Harries and Neil Symington

===================================================================
General comments

Introduction, 3rd paragraph: The blue-ward asymmetric emission is
generated in magnetospheric accretion models not only by the
occultation of the flow by the stellar photosphere, which occults
mostly high velocity blue-shifted material, but also by the
occultation of part of the receding flow by the inner disk, which
occults mostly low velocity redshifted emission. Both effects (lack of
low velocity red-shifted emission and lack of high velocity
blue-shifted emission) contribute to generate the blue-ward asymmetry.

   * Thank you for point that out. I replaced the sentence 
     from 
	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the stellar photosphere, .... 
     to

	  "... blue-ward asymmetric emission line 
	    profiles as resulting from the partial occultation 
	    of the flow by the inner part of the accretion disc and by
	    the stellar photosphere, .... "

Section 2.1, 1st paragraph: The adopted parameters of a typical
classical T Tauri star are listed in Table 1. I think that a mass
accretion rate of 10^-8 Msun/yr is much more representative of a
typical CTTS than 10^-7 Msun/yr. The values you adopted as typical for
both Macc and Mwind rates are actually quite high compared to what is
usually measured (see for ex. Gullbring et al. 1998). It is true that
some CTTSs present such high Macc and Mwind rate values, but these are
only a few. This is just a comment and does not invalidate your
results since you ended up calculating profiles for Macc = 10^-7,
10^-8 and 10^-9 Msun/yr, which clearly encompasses the whole range of
measured Macc rates in CTTSs.

    * I agree. A typical value of mass-accretion rate is ~3x10^-8
      Msun/yr according to Gullbring et al. (1998) and  Calvet et al. (2004).
      A slightly higher mass-accretion was used so that the line 
      broadening can be seen more clearly, and because 
      10^-7 is a nice round number. I clarify this point in caption of
      Table 1.

Section 2.1, 2nd paragraph: A better reference to the accretion schock
characteristics would be Calvet & Gullbring (1998) ApJ 509, 802 and
Gullbring et al. (2000) ApJ 544, 927, instead of Konigl (1991) and
Hartmann et al. (1994).

     * I replaced the references as you suggested.

In this paragraph you obtain about 8% of the stellar surface covered
by hot rings. This is quite a large value compared with the values
obtained by Gullbring & Calvet in the papers cited above, which range
rather from 0.1% to 1% for several CTTSs. But this could be due to the
fact that you are using a large Macc rate (10^-7 Msun/yr) for a
CTTS. Does this affect your results ? Was the hot ring surface
coverage recalculated consistently when you computed models with 10^-8
Msun/yr and 10^-9 Msun/yr ? In that case, what values were used ?

   * The surface "area" covered by the hot rings depends only on 
     the geometry of the magnetosphere (Rmi and Rmo); however, 
     the total accretion luminosity (L_acc) depends on the
     mass-accretion rate (Macc).
     The surface intensity of the hot rings is computed according to 
     this total luminosity due to accretion and by assuming they
     radiates as a blackbody.

     If a smaller hot ring surface was adopted, the ring temperature
     will increase in order to have a fixed L_acc for a given amount
     of mass-accretion rate.  So, yes, we calculate the amount of
     total "flux" according to a mass-accretion rate, but the surface
     coverage does not change.

     In our model, the hot ring simply acts
     as a source of additional continuum source, but the total amount
     of the additional continuum flux from the hot ring does not
     change.  This means that the level population does not change;
     hence, the H-alpha line strength is not affected by the change in
     the area coverage of the rings provided that the change is not so
     big. If the area coverage is much bigger (e.g. 30%), then the
     geometry of the magnetosphere itself changes significantly.  This
     would result in the line strength slightly for a fixed
     inclination angle, but overall shape of the profile would not
     change so much. For example, we used Rmi=2.2R* and Rmo = 3.0R*
     (Small and Wide (SW) model of Muzerolle et al 2001) which results
     in 8% hot ring coverage. If we were to use Rmi=2.2R* and Rmi =
     2.8R* and Rmo=3.0* (Small and narrow (SN) model of Muzerolle et
     al 2001), then the coverage would be 1% of the total surface,
     which is more similar to the fraction estimated in Gullring et al
     (2000).  If we look at the model comparison of Muzerolle et al
     (2001) (see their Fig. 5), the line strength changes, but the
     line profiles shape do not change much (except for the 10^3 K
     model). We confirmed, by running additional models, the line
     shapes do not change significantly for H-alpha by changing from
     SW to SN model.

     To make it clear about this point, I have inserted the following
     texts near the end of the 2nd paragraph in  Section 2.1:

        "In our model, the fraction of hot-ring coverage area is
        determined only by the geometry of the magnetosphere (Section
        2.2). Gullring et al (2000) estimated the fractions of the
        hot-ring coverage area for three intermediate CTTS by using
        the observed UV and optical spectra. They found the range of
        fraction to be 0.3--3.0~per~cent.  This is smaller than the
        value adopted here; however, even if we use 1~per~cent
        hot-ring area coverage model (by changing the width of the
        magnetosphere), we find the line shapes of H$\alpha$ do not
        change significantly. "

     Also I added the following sentence, just before the last
     sentence in Section 2.1:

        "The luminosity ratio and the temperature depend on a adopted
         mass-accretion rate."

Sect. 2.2, 3rd paragraph: You say that the results of Martin (1996)
qualitatively agree with those of Hartmann et al (1994). There is a
comment about that in the Muzerolle et al (1998) (ApJ 492, 743) paper
on page 751, where they also use the Hartmann et al (1994) law. They
say that the Martin (1996) temperature distribution is inadequate to
obtain line emission consistent with the observations, even at Halpha,
while their temperature distribution generates profiles that look like
the observed ones. So I think that you should not say that both
temperatures agree, even qualitatively.  It gives the idea that they
generate the same types of profiles, which is not true. The fact that
they do not agree is actually a big problem, since the Hartmann et
al. (1994) temperature structure, which is somewhat arbitrary,
reproduces the observed profiles, while the Martin (1996) temperature
structure, which was calculated consistently, does not. We are missing
some physics in there.

   * Thank you for reminding me this point. 
     I removed the sentence, "The results of Martin(1996)
     qualitatively agree with those of Hartmann et al. (1994)" from the
     text, and added the following at end of the paragraph to reflect
     your comments.

        "Although the model of Martin (1996) seems to be more
         comprehensive than that of Hartmann et al. (1994), Muzerolle
         et al (1998) found that the line profile models computed
         according to the temperature structure of the former do not
         agree with observations while they do have much better
         agreement when the temperature of the latter model was used. 
	 The temperature structure of the magnetosphere is still a big
         uncertainty in the accretion model; however, for a
         heuristic reason, we adopted the model of Hartmann et
         al. (1994). "

Sect. 2.3, 4th paragraph: I could not follow you in the end of the
paragraph. You have p=4*alpha*q and q = -1.15 from Whitney et
al. (2003). You say that in order to be consistent with Krasnopolsky
et al. (2003) who used p = -5/2, the value of alpha was set to
0.76. Well, alpha=p/(4*q)=-2.5/(4*(-1.15))=0.54 and not 0.76.

   * You are absolutely correct. This is my mistake.
     It should be p=-7/2 which is the value used in the models.  
     With p=-7/2,  
         alpha=p/(4*q)=-3.5/(4*(-1.15)) = 0.76.

	 ==> Corrected.

Sect. 4.1, 1st paragraph Tmax is not defined. I know it is the maximum
temperature of the accretion flow, but you must define it.

   * I added the following sentences after the first sentence of the
     same paragraph:
       "The temperature structure of Hartmann et al. (1994, ApJ, 426,
	669) is  adopted here, and the parameter Tmax sets
	the maximum temperature of the accretion streams. "

Sect. 4.1, 5th paragraph You say the EWs decrease as the inclination
increases mainly because the fraction of the accretion stream blocked
by the photosphere increases. I agree that the fraction of the
accretion stream blocked by the photosphere increases a lot as the
inclination increases (in Fig. 1, lower left funnel portion or z<0 and
x<0 for an observer looking from the upper right corner), but the
fraction blocked by the inner disk (at z<0 and x>0 in Fig. 1) also
increases a lot and contributes to the decrease in the EWs. The
photosphere mostly blocks blue-shifted material, while the inner disk
mostly blocks redshifted material.

   * You are absolutely right. I changed the sentence as you pointed
     out.

     ORIGINAL 
       "The line flux and the EWs decrease mainly because the fraction
        of the accretion stream blocked by the photosphere increases
        as the inclination increases."

     REVISED
       "The line flux and the EWs decrease mainly because (1) the
        fraction of the accretion stream blocked by the photosphere
        increases as the inclination increases (c.f. Fig. 1,
        lower-left part of the accretion stream i.e.  z<0 and x<0),
        and (2) the fraction of the accretion stream blocked by the
        inner part of the accretion disc increases as the inclination
        increases (c.f. Fig. 1, lower-right part of accretion stream
        i.e. z<0 and x>0). While the photosphere mostly blocks the
        blue-shifted part of the stream, the inner disc blocks the
        red-shifted part."

Sect. 4.1, 6th paragraph I don't understand why the highest
red-shifted line-of-sight velocity is visible only at the high
inclination angles. Looking at Figure 1, at low inclinations (even at
i=0) we actually see all the highest red-shifted material of the upper
right funnel portion (z>0, x>0, for an observer looking from the upper
right corner). Actually, looking at the profiles in Fig. 5, I don't
see much of a change in the redshifted wing width (for example at v >
120 km/s) with the increase in inclination. The blue-shifted wing
changes more.

   * Sorry. I meant to say "the blue-shifted" line-of-sight velocity
     instead of the red-shifted ...  As you mentioned the difference
     in the red wing is very small, but the difference in the blue
     wing is small but visible. By reflecting these points, I have
     made the correction as follows:

       ORIGINAL
       "..., the highest red-shifted line-of-sight velocity is visible
        only at the high inclination angles. This explains the wider
        appearance of the profile with i=80 deg. compared to the
        relatively narrow line appearance of the profile with
        i=10 deg."

       REVISED 
       "..., the highest blue-shifted line-of-sight velocity is seen
        only at the high inclination angles. For example, the
        accretion stream close to the photosphere (~1 R*) above the
        photosphere)on the upper-left quadrant in Fig.1 (x<0 and z>0)
        will be more aligned with the line-of-sight of an observer in
        the upper-right quadrant in the same figure (x>0 and z>0) as
        the inclination increases.  This results in a slight increase
        in the blue wing flux for a higher inclination model as seen
        in Fig. 5."

   * In addition to this correction, the last sentence in this
     paragraph (Sect. 4.1, 6th paragraph),

     "A similar EW dependence on
      inclination angle is found for the models with different 
      magnetospheric temperatures and the mass-accretion rates."

      was moved to the end of the previous paragraph (Sect. 4.1, 6th
      paragraph) because it was out of place.

Sect. 4.2, end of 2nd paragraph: You comment that the blue edge of the
absorption component in the slowest wind acceleration model (beta=4.0)
is located close to line centre (~20 km/s). But you don't show
beta=4.0 models. Was that a mistake and you intended to say beta=2.0 ?
I think you should cite the blue edge absorption component velocity of
the fastest model shown in Fig.6, which is beta=2.0. It is easier for
the reader to be able to check in the figure what you are explaining
in the text, and you make your point in the same way.

   * You are right again. I meant to say beta=2.0 instead of
     beta=4.0.  ==> Corrected.

Sect. 4.2, Table 4: Were does the value of p= -3/2 come from ? In
Section 2.3 you said you would be consistent with the model of
Krasnopolsky (2003) who used p= -5/2.

   * It should be p= -7/2.  
     A similar correction was made in Section 2.3 as you also pointed out.
     Sorry for the careless mistakes.
     ==> Corrected.

Sect. 4.2, 4th paragraph: The opening angle is measured from the
symmetry axis (z), then it should be given by theta_open=arctan(Rwi/d)
instead of theta_open=arctan(d/Rwi). High values of d correspond to
more collimated winds and therefore to smaller opening angles.

   * My mistake again. It should have written as
     theta_open=arctan(Rwi/d). 
     ==> Corrected.

Sect. 4.2, end of 4th paragraph: You say that you used d=44 Rstar for
the disc wind models presented in the paper but Table 4 has d= 22
Rstar listed. Which one is correct ? If Table 4 is a summary of the
reference disc wind models used in Section 4.2, maybe you should not
list d=22 Rstar, since you actually varied d in that section. Either
you put the range of values used (from 2 Rstar to 88 Rstar, as in
Fig. 7) or you don't put any value for d in Table 4.

   * This is my mistake again. I got confused between R* and Rsun.
     The largest displacement (d) model is "d=22R*." 
     d=44R* and d=88R*
     should be replaced with d = 22R*. "d=22R*" is the value used in 
     the models shown in Figure 6.
     ==> The values are corrected in the text and in the caption of
     Figure 7.

   * Deleted the entry for d from Table 4 as suggested.

Sect. 4.2: A general comment on the wind models presented in this
section. Blandford & Payne (1982) showed that in order to launch a
wind magneto centrifugally, the poloidal component of the magnetic
field must make an angle of less than 60 degrees with the disk
surface. In order for that to be achieved in your model, you must have
(supposing a flat disc, which should be reasonable for the inner disk
regions):

d < Rw*tg(theta), where theta = 90 - theta_open = 60 degrees is the
angle between the field line and the disk surface (x axis).

This means that for Rw = Rwi = 3 Rstar, you must have d < 5.2 Rstar
and that for Rw = 1 au ~ 100 Rstar, you must have d < 173 Rstar

The last condition is clearly satisfied in all your models, since d is
always less than 173 Rstar. But the inner field lines, passing close
to Rwi, do not satisfy the Blandford & Payne (1982) condition, since
for that you would need to have d < 5.2 Rstar in all cases.

In other words, if you fixed d = 44 Rstar in the disc wind models
presented in the paper, as you say you did in the 4th paragraph of
Sect. 4.2, for the wind to be launched centrifugally according to
Blandford & Payne, the field lines must present Rw > d/tan(60), which
is Rw > 25.4 Rstar. This means that field lines from Rwi = 3 Rstar to
Rw = 25 Rstar will not be able to launch a wind magneto-centrifugally
with d=44 Rstar.

How do you launch the wind in the inner field line region ? If you
cannot launch a wind in that region, how does it affect your results ?

   * I noticed the launching angle (measured from disc plane)
     condition of Blandford & Payne (1982) is not satisfied in the
     inner region of our simple disc wind model.  This is was done so
     that wind has a moderate amount of collimation. Since the stream
     line is straight, the wind will never bend toward the symmetry
     axis (z-axis) as the distance from the disc increases unlike the
     model of Blandford & Payne (1982). We therefore used a larger
     launching angle to simulate the stream lines in the far field (at
     a larger distance from the disc plane) in their model.  This is
     an intrinsic problem with approximating the stream lines with
     straight lines. A better model is perhaps to compute the field
     lines using Blandford & Payne (1982) model as done in
     e.g. Alencar et al (2005, A&A, 440, 595).  Also, in our mode, if
     the launching angle is too small (<60 deg.), the wind absorption
     feature would not been seen for profiles computed with i<60
     deg. since the wind stream line is straight and little material
     is present in the line-of-sight to an observer in this case.
     Since more than 50% of observed H-alpha profiles (e.g. Reipurth
     1996, A&AS, 120, 229 and Alencar & Basri 2000, ApJ 119, 1881) are
     affected by wind absorption, it was necessary for our model to
     used the launching angle (at the inner most part) greater than 60
     degrees to be more consistent with observation.

     If the launching angle (measured from the disc plane) is 60 deg. at
     the inner most part of the  wind, and the inclination angle
     (measured from the symmetry  axis, z-axis) is less than 60 deg., little
     absorption is expected, as demonstrated in Fig. 7.
     The situation will however change if the
     field/stream line is not straight and bending towards the
     symmetry axis as the wind travel higher above the disc.

     In summary, this was necessary compromise in order (1) to be more
     consistent with observation, and (2) to simulate more
     "colliminated" wind (toward the symmetry axis).

     To answer your question how it would effect our results, I run
     additional models with d=5Rstar which satisfies the wind
     launching condition of Blandford & Payne (1982) in all disc wind
     zone in our model.  These models showed that the wind absorption
     and emission features weakens,  but overall morphology dependence
     of the profiles on the wind temperature and the wind acceleration rate
     is very similar to what we found in the models shown in Figs. 6
     and 9.  Hence, the major conclusions obtained by using the wind
     models with a larger launching angle (at the very inner part of
     the disc) presented in this paper would not differ from the
     conclusions obtained from the disc wind models with d=5Rstar.

     To clarify the points you raised above, I have inserted the
     following in text (mostly the same as mentioned above).

     After the 4th paragraph in Section 4.3, the following new
     paragraphs are inserted:

       "Blandford & Payne (1982) showed that in order to launch a wind
        from a disc plane magneto-centrifugally, the launching angle
        (theta_L, measured from the equatorial disc plane) must be
        less than 60 degrees. In our models, the wind launching zone
        is restricted in between Rwi < wi < Rwo (see Fig. 1 and Table
        1) with Rwi = 3 Rstar and Rwo=1 au.  This sets the upper limit
        of the source displacement distance (d) to be around 5.2 Rstar
        if we want to satisfy the launching angle condition of
        Blandford & Payne (1982) for all the stream lines in our disc
        wind model.  In the models presented in Fig. 6 (and later in
        Fig. 9 in next section), we have adopted d =22 Rstar
        instead. This means that some part of the disc wind (3Rstar <
        wi < 25.4 Rstar) may not be able to be launched from the disc
        plane according to the model of Blandford & Payne (1982).

	We have used a larger launching angle or a larger source
	displacement distance for the following two reasons: (1) to
	simulate more 'colliminated' wind (toward the symmetry axis,
	and (2) to be more consistent with the fraction of H-alpha
	profiles affected by wind absorptions in observations.

### 2006-04-05 TTau_Halpha/figures/appedix.tex
Revised version after the refree's report
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@693 25464f21-6099-402c-b30a-640f1d40d38b

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}

\setcounter{tocdepth}{3}
\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Bold symbol macro for standard LaTeX users
\newcommand{\boldsymbol}[1]{\mbox{\boldmath $#1$}}

%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\newcommand{\aap}{A\&A}
\newcommand{\aaps}{A\&AS}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{\apj}
\newcommand{\pasp}{PASP}
\newcommand{\aj}{AJ}
\newcommand{\mnras}{MNRAS}
\newcommand{\apjs}{ApJS}
\newcommand{\aapr}{A\&AR}
\newcommand{\pasj}{PASJ}

\newcommand{\kmps}{\mathrm{km~s^{-1}}}
\newcommand\ion[2]{#1$\,${\sc {#2}}}   % ion, i.e., CII = \ion{C}{ii}
\newcommand{\Kelvin}{\mathrm{K}}

\usepackage{babel}
\makeatother
\begin{document}
\title[On the formation of H$\alpha$ around classical T Tauri
  stars]{On the formation of H$\alpha$ line emission around classical T Tauri stars}

\author[R. Kurosawa et\,al.]{Ryuichi
  Kurosawa$^1$\thanks{E-mail:rk@astro.ex.ac.uk}, Tim J. Harries$^1$
  and Neil H. Symington$^2$\\ $^1$School of Physics, University of Exeter, Stocker Road, Exeter EX4~4QL. \\ $^2$School of Physics and Astronomy, University of St. Andrews, North Haugh, St. Andrews, Fife, KY16 9SS.}

\date{Dates to be inserted}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2005}

\maketitle

\label{firstpage}

\begin{abstract}

We present radiative transfer models of the circumstellar environment
of classical T~Tauri stars, concentrating on the formation of the
H$\alpha$ emission. The wide variety of line profiles seen in
observations are indicative of both inflow and outflow, and we
therefore employ a circumstellar structure that includes both
magnetospheric accretion and a disc wind.  We perform
systematic investigations of the model parameters for the wind and the
magnetosphere to search for possible geometrical and physical
conditions which lead to the types of profiles seen in
observations. We find that the hybrid models can reproduce the wide
range profile types seen is observations, and that the most common
profile types observed occupy a large volume of parameter
space. Conversely, the most infrequently observed profile morphologies
require a very specific set of models parameters.  We investigate the
wind contribution to the line profile as a function of model parameters, and
examine the reliability of H$\alpha$ as a mass accretion diagnostic.
Finally, we examine the
H$\alpha$ spectroscopic classification used by Reipurth et.~al, and
discuss the basic physical conditions that are required to reproduce
the profiles in each classified type.

\end{abstract}

\begin{keywords}
stars: formation -- circumstellar matter -- radiative transfer -- stars: pre-main-sequence
\end{keywords}

\section{Introduction }

\label{sec:Introduction}

T~Tauri stars (TTS) are young ($\gtrsim 3\times10^{6}\,\mathrm{yrs}$,
\citealt{appenzeller:1989}) low-mass objects, and are the progenitors
of solar-type stars. Classical T~Tauri stars (CTTS) exhibit strong
H$\alpha$ emission, and typically have spectral types of F--K. Some of
the most active CTTS show emission in higher Balmer lines and metal
lines (e.g., \ion{Ca}{ii}~H and K). They also exhibit excess continuum
flux in the ultraviolet (UV) and infrared (IR).  Their spectral energy
distribution and polarisation data suggest the presence of
circumstellar discs, which plays an important role in regulating
dynamics of gas flows around CTTS (e.g. \citealt{camenzind:1990}).

Many observational studies (e.g., \citealt{herbig:1962};
\citealt{edwards:1994}; \citealt{kenyon:1994};
\citealt*{reipurth:1996}; \citealt{alencar:2000}) of CTTS line
profiles have revealed evidence for both outward wind flows and inward
accretion flows, characterised by the blue-shifted absorption features
in H$\alpha$ profiles and the redshifted inverse P~Cygni (IPC)
profiles respectively. Typical mass-loss rates of CTTS are about
$10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kuhi:1964}; \citealt{edwards:1987};
\citealt*{hartigan:1995}), and the mass-accretion rates are also about
$10^{-9}\,\mathrm{M_{\sun}\, yr^{-1}}$ to $10^{-7}\,\mathrm{M_{\sun}\,
yr^{-1}}$ (e.g., \citealt{kenyon:1987}; \citealt*{bertout:1988};
\citealt{gullbring:1998}).  Recent H$\alpha$ spectro-astrometric
observations by \citet*{takami:2003} provide indirect evidence for the
presence of bipolar and monopolar outflows down to $\sim
1\,\mathrm{au}$ scale (e.g.\,CS~Cha and RU~Lup).  Similarly, ESO VLT
observations using high-resolution ($R=50\,000$) two-dimensional
spectra of edge-on CTTS (HH30$^{*}$, HK~Tau~B, and HV~Tau~C) by
\citet{appenzeller:2005} show extended H$\alpha$ emission in the
direction perpendicular to the obscuring circumstellar disc,
suggesting the presence of the bipolar outflows. On an even larger
scale, \emph{HST} observations of HH30 \citep{burrows:1996} trace the
jet to within $\lesssim30$~au of the star. The jet has a cone shape
with an opening angle of $3^{\circ}$ between 70 and 700 au
\citep{koenigl:2000}. \citet{alencar:2000} found about 80 per cent of
their sample (30 CTTS) show blue-shifted absorption components in at
least one of the Balmer lines and Ca~K.

In the currently favoured model of accretion in CTTS, the accretion
disc is disrupted by the magnetosphere, which channels the gas from
the disc onto the stellar surface (e.g., \citealt{uchida:1985};
\citealt{koenigl:1991}; \citealt{cameron:1993};
\citealt{shu:1994}). This picture is supported by recent measurements
of strong ($\sim 10^{3}\mathrm{G}$) magnetic fields in CTTS (e.g.,
\citealt{johns-krull:1999}; \citealt{symington:2005b}) and by
radiative transfer models which reproduce the gross characteristics of
observed profiles for some TTS (\citealt*{muzerolle:2001}). In
particular the magnetospheric accretion (MA) model explains blue-ward
asymmetric emission line profiles as resulting from the partial
occultation of the flow by the stellar photosphere, while inverse
P~Cygni profiles in the MA model result from inflowing material at
near free-fall velocities seen projected against hotspots on the
stellar surface.

Despite these successes, the overwhelming observational evidence for
outflow in the CTTS suggests that the MA model is only one component
of a complex circumstellar environment. Clearly, one must include the
contribution of any wind/jet flow if one wishes to both accurately
predict the mass-accretion rate and also determined the mass-loss rate
of CTTS via emission profile modelling. The first attempt in this
direction was made by \citet{alencar:2005} who demonstrated that the
observed H$\alpha$, H$\beta$ and Na~D lines of RW Aur are better
reproduced by the radiative transfer model which included a collimated
disc-wind arising from near the inner edge of the accretion disc.

The main aim of this paper is to find a simple kinematic model which
can reproduce the wide variety of the observed profiles, and to
perform empirical studies of line formation in an attempt to place
morphological classification schemes on a firmer physical footing. We
will also discuss whether our model is consistent with some
predictions made by recent (magneto-hydrodynamics) MHD studies i.e.~
$\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}\approx0.1$
(e.g.~\citealt{koenigl:2000}).

In section~\ref{sec:model-configuration}, the model assumptions, and
the basic model configurations are presented. We discuss the radiative
transfer model used to compute the profiles in
section~\ref{sec:radiative-transfer-model}, and the results of model
calculations are given in section~\ref{sec:Results}.  We discuss our
results in the context of Reipurth's classification scheme in
section~\ref{sec:Discussion}, and our summary and conclusions are
presented in section~\ref{sec:Conclusions}.

%%%%%%%%%%%%%%

\section{Model configuration}

\label{sec:model-configuration}

In order to understand how the different parts of the CTTS
circumstellar environment contribute to the formation of H$\alpha$,
the model space is divided into four different regions: (1)~a central
continuum source, (2)~the magnetospheric accretion flow, (3)~the wind
outflow, and (4)~the accretion disc. Fig.~\ref{fig:config_discwind}
depicts the relative location of the regions in the model space.  The
density is assumed to be rotationally symmetric around the
$z$-axis. The innermost radius of the magnetosphere at the equatorial
plane coincides with the the inner radius of the accretion disc. From
the innermost part of the accretion disc, the gas falls freely, moving
along the magnetic field onto the surface of the star.  In the
following subsections we describe the details of model components.

\subsection{The continuum source}

\label{sub:Continuum-Source}

We adopt stellar parameters of a typical
classical T~Tauri star for the central continuum source, i.e. radius
($R_{*}$), mass ($M_{*}$), and effective temperature the photosphere
($T_{\mathrm{ph}}$) are $2\, R_{\sun}$, $0.5\, M_{\sun}$, and $4000\,\mathrm{K}$
respectively. The model atmosphere of \citet{kurucz:1979} with
$T_{\mathrm{ph}}=4000\,\mathrm{K}$ and $\log g_{*}=3.5$ (cgs) defines
the photospheric contribution to the continuum flux. The
parameters are summarised in Table~\ref{tab:std_parameters}.

An additional continuum source is considered for models which includes
magnetospheric accretion: as the infalling gas approaches the stellar
surface, it decelerates in a strong shock, and is heated to $\sim
10^{6}\,\mathrm{K}$.  The X-ray radiation produced in the shock will
be absorbed by the gas locally, and re-emitted in optical and UV
radiation (\citealt{koenigl:1991}; \citealt*{hartmann:1994}). This
will create hot rings where the magnetic field intersects with the
surface. We assume that the free-falling kinetic energy is thermalized
in the radiating layer, and is re-emitted as blackbody radiation with
a single temperature. With the parameters of the magnetosphere and the
star given above (Table~\ref{tab:std_parameters}), about $8$~per~cent
of the surface is covered by the hot rings. If the mass-accretion rate
is $10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, the ratio of this accretion
luminosity to the photospheric luminosity is about $0.5$, and the
corresponding temperature of the hot rings is about
$6400\,\mathrm{K}$.  The continuum emission from the hot rings is
taken into account when computing the line profiles.

\begin{table*}

\caption{Summary of the reference classical T~Tauri star model parameters.}
\label{tab:std_parameters}

\begin{center}

\begin{tabular}{lccccccccccc}
\hline 
Parameters&
$R_{*}$&
$M_{*}$&
$T_{\mathrm{ph}}$&
$R_{\mathrm{mi}}$&
$R_{\mathrm{mo}}$&
$\dot{M}_{\mathrm{acc}}$&
$\dot{M}_{\mathrm{wind}}$&
$v_{\infty}$&
$b$&
$R_{\mathrm{di}}$&
$R_{\mathrm{do}}$\tabularnewline
&
$\left[R_{\sun}\right]$&
$\left[M_{\sun}\right]$&
$\left[\mathrm{K}\right]$&
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[\mathrm{km\, s^{-1}}]$&
$[-]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$\tabularnewline
\hline
&
$2.0$&
$0.5$&
$4000$&
$2.2$&
$3.0$&
$10^{-7}$&
$10^{-8}$&
$210$&
$4.0$&
$2.2$&
$100$\tabularnewline
\hline
\end{tabular}

\end{center}

\end{table*}

\subsection{The magnetosphere}

\label{sub:Magnetosphere}

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_dipolar}, the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

The magnetic field and the gas stream lines are assumed to have the
following simple form:
\begin{equation}
      r=R_{\mathrm{m}}\sin^{2}\theta
  \label{eq:dipole}
\end{equation}
\citep*[see][]{ghosh:1977} where $r$, and $\theta$ are coordinates
of the field point ($p$) in Fig.~\ref{fig:config_dipolar} in spherical
coordinates, and $R_{\mathrm{m}}$ is the radial distance to the field
line at the equatorial plane ($\theta=\pi/2$). The range of
$R_{\mathrm{m}}$ is restricted between $R_{\mathrm{mi}}$ and
$R_{\mathrm{mo}}$. Using the field geometry above and conservation of
energy, the velocity and the density of the accreting gas along the
steam line are found as in \citet{hartmann:1994}.

The temperature structure of the magnetospheric used by
\citet{hartmann:1994} is adopted here. They computed the temperature,
assuming a volumetric heating rate which is proportional to $r^{-3}$,
by solving the energy balance of the radiative cooling rate of
\citet{hartmann:1982} and the heating rate \citep{hartmann:1994}.
\citet{martin:1996} presented a self-consistent determination of the
thermal structure of the inflowing gas along the dipole magnetic field
(equation~\ref{eq:dipole}) by solving the heat equation coupled to the
rate equations for hydrogen. He found that main heat source is
adiabatic compression due to the converging nature of the flow, and
the major contributors to the cooling process are bremsstrahlung
radiation and line emission from Ca~II and Mg~II ions. The results of
\citet{martin:1996} qualitatively agree with that of
\citet{hartmann:1994}.

\subsection{The disc-wind outflow geometru}
\label{sub:wind_jet}

The magneto-centrifugal wind paradigm, first proposed by
\citet{blandford:1982}, has been often used to model the large-scale
wind structure of T Tauri stars, or the observed optical jets
(e.g.~HH~30 jet by \citealt{burrows:1996}; \citealt{ray:1996}). The
launching of the wind from a Keplerian disc is typically done by
treating the equatorial plane of the disc as a mass-injecting boundary
condition (e.g., \citealt{shu:1994}; \citealt{ustyugova:1995};
\citealt{ouyed:1997}; \citealt{krasnopolsky:2003}).  Depending on the
location of the open magnetic fields anchored to the disc, two
different types of winds are produced. If the field is constrained to
be near the co-rotation radius of stellar magnetosphere, an
{}``X-wind'' \citep{shu:1994} is produced. If the open field lines are
located in a wider area of the disc, a {}``disc-wind'' similar to that
of \citet{koenigl:2000} is produced \citep{krasnopolsky:2003}.  Recent
reviews on the jet/wind-disc connection can be found in
\citet{koenigl:2000} and \citet{pudritz:2005}. Clearly there are several alternative outflow scenarios. Here we adopt
a simple kinematical model, based on the disc-wind paradigm, which broadly represents the results of
MHD simulations,

>>>>

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{graphicx}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\usepackage{babel}
\makeatother
\begin{document}
\begin{tabular}{c}
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc10.eps}\tabularnewline
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc55.eps}\tabularnewline
\includegraphics[%
  scale=0.45]{/usr/local/home/rk/paper/TTau_Halpha/figures/discwind_hybrid_mdot7_inc80.eps}\tabularnewline
\end{tabular}
\end{document}
